Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-10 paper.tex
Section 2

Timed automata targets real-time systems and employs finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories (behaviors) with constant slope. Hybrid automata bridges digital processing and analog entities and consequently generalizes timed automata, by also employing finite automata with real variables, whose trajectories follow more general dynamical laws~\cite{Henzinger95}.  Timed automata and hybrid automata are deterministic models for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}.

Finally, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. Generally, the properties that must hold for ECPSs are related to safety, liveness, and reachability. In particular, model-checking procedures have been employed to verify those properties even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking~\cite{Veanes09}.

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security. Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16} \textcolor{red}{eh bom citar mais fontes}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.

@@

@@

\begin{thebibliography}{1}

@@

\end{thebibliography}

>>>>

Timed automata targets real-time systems and employs finite automata with clocks, {\it i.e.}, real variables that follow continuous trajectories (behaviors) with constant slope. Hybrid automata bridges digital processing and analog entities and consequently generalizes timed automata, by also employing finite automata with real variables, whose trajectories follow more general dynamical laws~\cite{Henzinger95}.  Timed automata and hybrid automata are deterministic models for hybrid systems that built the basis of HS representations, which were subdivided into different classes of hybrid automata, depending on mapping functions ($r$ and $v$), {\it e.g.}, linear hybrid automata~\cite{Lafferriere99}, rectangular hybrid automata~\cite{Puri94,Henzinger95}, nonlinear hybrid automata~\cite{Henzinger95NL,Broucke1997}, integration graphs~\cite{Kesten1993}, and linear stochastic hybrid automata~\cite{Julius09}. Recently, Rungger and Tabuada~\cite{Rungger16} and Tabuada {\it et al.}~\cite{Tabuada14} proposed a general and wide model for ECPSs, which is based on hybrid automata, by providing the concept of robustness and important tools for robust stability analysis. In addition to HA and TA based models, there are a number of other ECPS modeling initiatives based on (exact or approximate) bisimulation~\cite{Girard11}, simulation~\cite{Rungger13} and interval analysis~\cite{Li16}.

Another approach to ECPS modeling concerns quantized control systems and finite-word length (FWL) effects on the performance of these systems. Differently from bisimulations, quantized models are built from a countable subset of the (quantized) input space~\cite{Tabuada07}. In particular, Keel and Bhattacharyya showed that even the robust and optimal controllers may still present an undesired property~\cite{Keel01,bhattacharyya}: fragility or sensitivity to implementation aspects ({\it e.g.}, the FWL representation of the digital controllers).  Therefore, various works tried link sampled-data theory to hybrid systems models~\cite{Bicchi02,Petreczky10,Petreczky07,Ye98}.

Finally, in addition to the modeling of the ECPS dynamics itself, it is often interesting to establish invariants on the state space of the ECPS, such information can be used in order to accelerate formal verification and synthesis processes for these systems. These invariants might be synthesized in order to point the boundaries of state-space~\cite{BenSassi12}, determine stability~\cite{Paul14}, or reachability of ECPS~\cite{Li16_2,Li16}.

@@

Given a class of ECPSs $\Omega$ and a set $\Phi$ of desired properties, a class of verification problems is called decidable if there is an algorithm that is able to decide if a finite number of steps $\forall \mathcal{O}\in\Omega$ satisfy every property $\mathcal{L}\in \Phi$~\cite{Alur00}. Decidability is a key issue in ECPS verification, since there are uncountable states and complex dynamics to be considered, {\it i.e.}, hybrid automata and their variations (usually employed to model ECPSs) are infinite-state machines~\cite{Henzinger95}. In particular, the safety verification of ECPS is undecidable, except in very specific cases as in systems that can be represented by timed automata and rectangular automata~\cite{Alur11}. An alternative to ensure that ECPS trajectories will not pass through unsafe states is the use of barrier certificates~\cite{Prajna07,Prajna04}. An example of barrier certificate is the stability, that ensures the convergence of states to an equilibrium point and consequently that there exists a region from where any trajectory remains inside of this region. The stability may be ensured based on Lyapunov theory (for a wide class of model abstractions)~\cite{tabuada2009verification}, or linear system stability theory for discrete simulations~\cite{Bessa17}.

Generally, the properties that must hold for ECPSs are related to safety/reachability, and liveness. Therefore, model-checking procedures have been employed to verify those properties even under undecidability conditions via approximations, {\it e.g.}, bounding the number of steps from the initial state, which leads to bounded model checking (BMC)~\cite{Veanes09}.

Along with the CPS modeling and verification techniques purely based on HA, appeared logic-based approaches~\cite{Phan15} that may still use HA, but it is not based on a algebraic description of states and transitions. Some examples of logic-based verification are the theorem proving techniques, {\it e.g.} deductive verification (i.e., interactive theorem proving)~\cite{Nipkow02} and automatic theorem proving~\cite{Platzer07}, and reduction-based techniques, among them the BMC stands out~\cite{handbook09}.

%Much of the work related to the formal verification of ECPS seeks to ensure the safety, robustness, and reachability properties of an HA. 
Recently, some of these works addressed the ECPS formal verification considering the discretization, switching and quantization effects.  Dugirala and Viswanathan~\cite{Duggirala2015} verified embedded control softwares considering real time issues, e.g., delayed responses and possible loss of real time deadlines. Anta {\it et al.}~\cite{Anta10} presented a tool called Costan that finds errors in the implementation of a mathematical model and verifies whether the error is tolerated, considering the quantization effects and the fixed-point implementation focusing on the quantization error effects over stability of the system. Similarly, Ismail {\it et al.}~\cite{dsv_spin2015} presets the DSVerifier, a BMC tool for digital systems that employs SMT/SAT solvers as back-ends in order to provide support for digital system design and verification considering FWL effects. In particular, DSVerifier verifies if a designed digital controller and/or closed-loop ECPS presents the desired performance when it is implemented in a FWL format~\cite{Bessa16,Bessa17}.

Finally, it is worth to mention that there are some other alternatives for modeling and analysis of ECPS, that are not based on verification, but in simulators~\cite{Phan2014,Nakajima12,Junjie12}. These approaches fails to ensure that some properties are hold but it presents a good scalability, what is ideal for large scale and complex systems that are hardly modeled through the automata- or logig-based approaches. Modelica~\cite{Junjie12} and Simulink/MATLAB~\cite{Nakajima12} are important examples of widely used simulators for CPS. In order to achieve a trade off between the scalability of CPS simulators and the correctness of verification methods, the statistical model checking have been employed to formally verify ECPS~\cite{Clarke11}.

In addition, Shoukry {\it et al.} \cite{Shoukry} tackled another important problem related to ECPSs: security. Indeed, the essence of ECPSs claims for sophisticated interfacing ({\it e.g.}, sensors and network connections), which normally leads to security issues~\cite{cpsattack}, given that wrong control actions can be generated through corrupted measurements. The mentioned authors proposed a methodology for estimating the state of physical systems, through formal methods, even if sensor inputs are contaminated with erroneous or intentionally corrupted data, in order to still be able to execute suitable control commands. The cyber-security is also a concern of Fiore {\it et al.} that aims to ensure the secure state estimation of UAV systems under sparse attacks~\cite{Fiore17}. Indeed, many situations can be regarded as satisfiability problems and solved, for instance, with approaches based on Satisfiability Modulo Theories (SMT)~\cite{Araujo16,TrindadeDAES2016,Rahman13}, which indicates many links with other research areas and might lead to a unified approach regarding system verification, security enforcement, and interaction handling.

@@

The control synthesis for ECPS consists of finding an controller or an invariant control law $\mathcal{C}$, such that a closed-loop system composed of $(\mathcal{C},\Omega)$ satisfies every property $\mathcal{L}\in \Phi$.

An efficient way to synthesize reliable ECPS is the Counterexample Guided Inductive Synthesis (CEGIS)~\cite{David15}. The key idea is the definition (by the user) of the safe states set, {\it i.e.} the specification of $\Phi$, as well as a generic template for a control system that must satisfy these specification; the the CEGIS engine compute values for the open values in the template that hold $\Phi$~\cite{RienerKFB16}. Abate~{\it et al.}~\cite{Abate17,abatecav2017} presents a method for synthesizing stable controllers that are suitable to continuous plants given as transfer functions~\cite{Abate17} and state-space representations~\cite{abatecav2017}, which exploits bit-accurate verification of software implemented in digital microcontrollers~\cite{Bessa16,Bessa17,dsv_spin2015}.

% The mentioned authors developed a tool called DSSynth, which is able to synthesize digital controllers that are algorithmically and numerically sound, w.r.t. stability. DSSynth marks the first use of counterexample-guided inductive synthesis~\cite{sketch} to synthesize digital controllers, considering physical plants with uncertain models and FWL effects; however, low-level implementation errors ({\it e.g.}, limit cycles) are not further investigated.

@@

%\begin{thebibliography}{1}

@@

%\end{thebibliography}
